%&LaTeX
% $Id: changes.tex,v 1.6 2001-06-04 01:32:52 paklein Exp $

\section{History of updates to the \tahoe User Guide}

\subsection{History}
\todo{
brief history:\\
originated by Stephanie Wimmer\\
updated to Version 3.4 by Patrick Klein.
}
\subsection{Changes since Version 3.4}
\label{sect.changes.log}
This section summarizes changes to the \tahoe input format
since Version 3.4.
\subsubsection{update 3.4.1}
Two additional output variables have been added to the element output
for solid deformation elements listed in
Table~\ref{tab.solid.output.element}. This brings the number of
element output flags from 5 to 7. This change is version aware. Only 5
flags are read if the input file version is older than 3.4.1.

\subsection{Modernisation track (post-3.4.1, GitHub)}
\label{sect.changes.modern}
After the move to GitHub (see Section~\ref{sect.cmake.build}), the
following capability changes have landed on top of the v3.4.1 input
format.  The format itself remains backwards compatible; the items
below describe new \emph{features}, new \emph{elements}, new
\emph{benchmarks}, and infrastructure cleanups that are not yet
folded into the per-section reference material elsewhere in this
guide.  Pull-request numbers refer to the GitHub tracker.

\paragraph{Build system.}
The legacy \texttt{tahoe-manager} Perl driver and CVS/SourceForge
workflow have been removed.  The project now builds with CMake (root,
\texttt{toolbox/}, \texttt{tahoe/}, plus per-library
\texttt{CMakeLists.txt} files); all bundled third-party C libraries
(\texttt{expat}, \texttt{SPOOLES}, \texttt{f2c}) are compiled from
source under \texttt{third\_party/} (PR \#44).  See
Section~\ref{sect.cmake.build} for build invocations and
Section~\ref{sect.platforms} for supported platforms.

\paragraph{Activated Drucker--Prager plasticity (Material 11).}
The small-deformation Drucker--Prager constitutive model is now wired
into the active element-type registry and verified on the
level.1 material suite (PR \#37).  The mathematical formulation in
Section~``Material 11'' of this guide already documented the model;
it is now available in stock builds without re-enabling.

\paragraph{Implicit Coulomb friction in \texttt{PenaltyContact3DT}.}
The 3D penalty-contact element now contributes a consistent tangent
contribution for Coulomb friction in implicit analyses (PR \#40 /
\#45).  Previously, only the residual was assembled, which forced the
nonlinear solver to fall back on a linear convergence rate.  With the
LHS contribution in place, frictional contact problems now converge
quadratically in Newton.

\paragraph{BonetTet element with lagged-J-bar and FD tangent.}
A new finite-deformation Tet4 element \texttt{<bonet\_tet>} has been
added, following the F-bar / ANP averaging of Bonet \etal\ (PR \#29).
The implementation caches J-bar at step start and builds the local
12$\times$12 tangent through a forward finite-difference of the F-bar
residual.  On moderately near-incompressible cases ($\kappa/\mu \sim
2.5$), Newton converges quadratically in $\sim$1 iter/step --- a
fifteenfold iteration reduction relative to the linear-rate baseline.
A fully consistent analytical tangent (which would handle the
cross-element $\partial \bar{J} / \partial \mathbf{u}$ coupling
required at severe near-incompressibility) is tracked as future work.

\paragraph{New benchmarks.}
The following benchmarks have been added to
\texttt{benchmark\_XML/level.5/} (PRs \#47, \#48, \#49 and the broader
explicit-solver track):
\begin{itemize}
\item \texttt{brinell/} --- Brinell indentation of a rigid sphere on a
$J_{2}$-plastic Hex8 block, exercising contact + plasticity together.
\item \texttt{tensile/} --- Quarter-symmetry uniaxial tension on a slim
Hex8 bar; isolates the $J_{2}$ radial-return from contact and verifies
that the implementation tracks an input \texttt{cubic\_spline}
$\sigma_{y}(\varepsilon^{p})$ hardening curve within 2\% under
monotonic loading.
\item \texttt{implicit\_friction/} --- Companion regression case for
the new implicit Coulomb friction tangent (PR \#40 / \#45).
\item \texttt{hertz/} --- Hertz contact validation case, including a
mixed Tet4 indenter on Hex8 base variant for the Tet element track.
\item \texttt{tet\_classic/} --- Classic-Tahoe implicit Tet4 reference
for the BonetTet element (PR \#29).
\end{itemize}
All per-level benchmark directories now carry a \texttt{README.md}
that catalogues the input files and references the originating issue
(PR \#42).

\paragraph{Documentation infrastructure.}
The User Guide source previously depended on \texttt{hangcaption.sty},
which was dropped from modern \TeX{}~Live.  The \texttt{\textbackslash hangcaption}
macro is now aliased to \texttt{\textbackslash caption} in
\texttt{tahoe.user.tex}; modern \texttt{\textbackslash caption} already
hangs the label.  No figure layouts change as a result.