%&LaTeX
% $Id: regression.tex,v 1.4 2026-05-23 saman Exp $

\section{Regression tests and benchmarks}
\label{sect.regression}
The repository ships a layered benchmark suite under
\texttt{benchmark\_XML/} that doubles as a regression catalogue and a
collection of fully-worked input examples.  Each level has a
\texttt{run.batch} that drives the bundled \texttt{tahoe} binary
through every input file in that level, and a companion
\texttt{compare} tool diffs the produced results against committed
reference data.  The complete entry point is \texttt{benchmark\_XML/ReadMe},
and each level has its own \texttt{README.md} describing the cases in
that tier.

\subsection{Benchmark levels}
\label{sect.benchmark.levels}
\begin{description}
\item[Level 0 --- Core capabilities.]
Smoke tests for the springs, solid mechanics, diffusion, particle,
meshfree, and time-integrator paths.  All code changes are expected to
keep level.0 green before being merged.
\item[Level 1 --- Constitutive verification.]
Bulk material models, cohesive surface relations, and atomistic
potentials exercised under simple loadings whose closed-form or
high-quality reference answers are known.
\item[Level 2 --- Extended capabilities.]
Larger models and longer run times than level.1.  Still serial and
self-contained (no optional modules required): $K$-fields, contact,
conveyor, thermostats, tied interfaces, torsion, controllers.
\item[Level 3 --- Parallel.]
MPI distributed-memory tests.  Requires building \tahoe with
\texttt{-DTAHOE\_ENABLE\_MPI=ON} and launching through
\texttt{mpirun -np 4}.
\item[Level 4 --- Solver performance.]
Multi-physics performance cases used to compare sparse direct solvers
(SPOOLES vs MUMPS vs SuperLU) on non-trivial DOF counts.  The
representative case is a 2D dielectric elastomer with a liquid
inclusion (\texttt{dielectric\_elastomer\_Q1P0Elastocapillary}, $\sim$18k DOF).
\item[Level 5 --- Explicit-solver and modern element benchmarks.]
Benchmarks added alongside the modernised explicit solver and the new
finite-deformation elements.  Includes the original
legacy-vs-vectorized speed-up matrix
(\texttt{explicit\_benchmark/}), the classic Taylor bar
(\texttt{taylor\_bar/}), the Hertz contact case (\texttt{hertz/})
covering both explicit and implicit pathways and a mixed Tet4/Hex8
indenter variant, the implicit-friction case
(\texttt{implicit\_friction/}, see~\ref{sect.changes.modern}), the
$J_{2}$-plasticity Brinell indentation
(\texttt{brinell/}, see~\ref{sect.changes.modern}), the uniaxial
tension verification of cubic-spline isotropic hardening
(\texttt{tensile/}, see~\ref{sect.changes.modern}), and the
classic-Tahoe BonetTet implicit Tet4 path
(\texttt{tet\_classic/}, see~\ref{sect.changes.modern}).
\end{description}

\subsection{Running a level}
From the repository root:
\begin{center}
\begin{minipage}{0.85\linewidth}
\begin{alltt}
\$ cd benchmark_XML/level.0
\$ printf "run.batch\(\backslash\)nquit\(\backslash\)n" | ../../build/bin/tahoe
\end{alltt}
\end{minipage}
\end{center}
Output files (\texttt{*.run}, \texttt{*.geo}, \ldots) are written
alongside the input XML files and are excluded from version control
via the project \texttt{.gitignore}.  A driver script
\texttt{run\_benchmarks.sh} at the repository root sweeps any subset of
levels --- e.g.\ \texttt{./run\_benchmarks.sh level.3} for the MPI
tier alone.  For level.3, the system \texttt{/usr/bin/mpirun} (Open MPI)
must be used; a Conda-shipped \texttt{mpirun} is typically MPICH and
ABI-incompatible with the OpenMPI that \tahoe links against.

\subsection{Repeatability of tests across platforms}
\label{sect.repeat.platforms}
The bundled \texttt{compare} tool (built under \texttt{build/bin/compare})
performs the file-by-file diff against the reference results
committed under each benchmark directory.  Cross-platform
discrepancies should be reported as issues on the project GitHub
tracker; numerical differences within the \texttt{compare} tolerance
are expected, while structural differences (missing fields, mismatched
mesh ordering) indicate a regression.

\subsection{Repeatability of tests during parallel execution}
\label{sect.repeat.parallel}
The level.3 tier is the authoritative parallel regression check.  A
test that passes serially but fails in parallel typically indicates a
problem in the partitioning of state variables across the partition
boundary, in the assembly of the global RHS, or in the
solver-specific communication path.
